---
title: "Analyse Biological Data"
author: "Gerard Temprano Sagrera"
date: "17/01/2022"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo = FALSE}
library(openxlsx);library(ggplot2);library(corrplot);library(ggpubr);library(corrr);library(ggbreak)
```

```{r}
bio <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/20221115_Datos Pacientes.xlsx")
bio <- bio[2:nrow(bio),]
bio$SEXO <- ifelse(bio$SEXO == 1, "Man","Woman")
```

```{r}
cases <- bio[bio$Status == "Case",]
controls <- bio[bio$Status == "Control",]
```

```{r}
cases2 <- cases[,7:ncol(cases)]
```

```{r}
#cases2[rowSums(is.na(cases2)) > 6,]
```

```{r}
#colSums(is.na(bio))
```

1.  Only sample, status, sex and age have no NAs.
2.  Cases only: Type, Symptomatic, Height, Weight, PAD, Other Aneurysm, Aortic Diameter, Leucocytes, Neutrophiles, Limphocytes, Monocytes, Platalets, Hb, RDW.
3.  Controls only: Pathology.

Which column would be word to try to complete?


## Correlation study:

### All:

```{r}
# Select columns:
bio2 <- bio[,2:ncol(bio)]
bio2 <- bio2[,c("Status","SEXO","age","TIPOANEURISMA","symptomatic","Height_cm","Weight_Kg","Smoking","other_aneurysm",
                "aortc_diameter_mm")]

# Transform to numeric:
bio2$Status <- as.numeric(as.factor(bio2$Status))
bio2$SEXO <- as.numeric(as.factor(bio2$SEXO))

# Remove pathology:
bio2 <- bio2[ ,!names(bio2) %in% c("Pathology")]

bio2 <- as.data.frame(lapply(bio2, as.numeric))
```


```{r, warning=FALSE}
M <- cor(bio2, use = "pairwise.complete.obs")
testRes <- cor.mtest(bio2, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

pdf(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correaltion_Plot_All.pdf")
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
dev.off()

corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
```

- Type of aneurysm, height, weight, other aneurysm and size are only available for cases.

- Status with Sex: There are more male among cases.

```{r}
tab <- as.data.frame(prop.table(table(bio2$Status, bio2$SEXO)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Status", fill = "Sex") +
  scale_x_discrete(breaks = c(1,2), labels = c("Case","Control")) +
  scale_fill_discrete(breaks = c(1,2), labels = c("Male","Female"))
```

- Status with Age: Cases are older.

```{r}
ggboxplot(bio2, x = "Status", y = "age") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 100) +
  scale_x_discrete(breaks = c(1,2), labels = c("Case","Control"))
```


- Status with Smoking: There are less smokers in controls.

```{r}
tab <- as.data.frame(prop.table(table(bio2$Status, bio2$Smoking)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Status", fill = "Smoking") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Case","Control")) +
  scale_fill_discrete(breaks = c(0,1,2), labels = c("Never","Past","Currently"))
```


- Sex with Smoking: There are more smokers among men (There are more cases in men).

```{r}
tab <- as.data.frame(prop.table(table(bio2$SEXO, bio2$Smoking)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Sex", fill = "Smoking Status") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Man","Woman")) +
  scale_fill_discrete(breaks = c("0","1","2"), labels = c("Never","Past","Currently"))
```

- Age with Weight: Weight reduced with age.

```{r, warning=FALSE}
ggscatter(bio2, x = "age", y = "Weight_Kg",
          add = "reg.line") +
  stat_cor(method = "pearson")
```

- Type of Aneurysm with Symptoms: There are more symptoms with Atherosclerotic aneurysms.

```{r}
tab <- as.data.frame(prop.table(table(bio2$TIPOANEURISMA, bio2$symptomatic)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Type", fill = "Symptoms") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Atherosclerotic","Inflammatory")) +
  scale_fill_discrete(breaks = c("0","1"), labels = c("Yes", "No"))
```


- Type with Aortic Diameter: Inflammatory aneurysms have bigger diameters.

```{r}
ggboxplot(subset(bio2, !is.na(TIPOANEURISMA)), x = "TIPOANEURISMA", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(1,2), labels = c("Atherosclerotic","Inflammatory"))
```


- Symptoms with Smoking:

```{r}
tab <- as.data.frame(prop.table(table(bio2$symptomatic, bio2$Smoking)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Symptomatic", fill = "Smoking Status") +
  scale_x_discrete(breaks = c("0","1"), labels = c("Asymptomatic","Symptomatic")) +
  scale_fill_discrete(breaks = c("0","1","2"),labels = c("Never", "Past", "Present"))
  
```


- Symptoms with Diameter: Symptomatic aneurysms are bigger.

```{r}
ggboxplot(subset(bio2, !is.na(symptomatic)), x = "symptomatic", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(0,1), labels = c("Asymptomatic","Symptomatic"))
```


- Weight with Height.

```{r, warning=FALSE}
ggscatter(bio2, x = "Weight_Kg", y = "Height_cm",
          add = "reg.line") +
  stat_cor(method = "pearson")
```


## Cases:

```{r}
cases <- bio2[1:97,]
cases <- cases[,2:ncol(cases)]
```

```{r}
M <- cor(cases, use = "pairwise.complete.obs")
testRes <- cor.mtest(cases, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

pdf(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correaltion_Plot_Cases.pdf")
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
dev.off()

corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
```

- Age with Weight: Weight reduces with age.

```{r, warning=FALSE}
ggscatter(cases, x = "age", y = "Weight_Kg",
          add = "reg.line") +
  stat_cor(method = "pearson")
```


- Age with Smoking: Currently smokers are younger.

```{r}
my_comparisons <- list( c("0","1"), c("0", "2"), c("1", "2") )

ggboxplot(subset(cases, !is.na(Smoking)), x = "Smoking", y = "age", color = "Smoking") +
  stat_compare_means(comparisons = my_comparisons, label.y = c(90, 110, 130), method = "t.test") +
  stat_compare_means(label.y = c(150), method = "anova") +
  scale_x_discrete(breaks = c("0","1","2"), labels = c("Never","Past","Currently Smoking"))
```


- Type of Aneurysm with Symptoms: Atherosclerotic aneurysms have more Symptoms.

```{r}
tab <- as.data.frame(prop.table(table(cases$symptomatic, cases$TIPOANEURISMA)))

ggplot(tab, aes(x = Var2, y = Freq, fill=Var1)) + geom_col() + labs(fill = "Symptomatic", x = "Type") +
  scale_fill_discrete(labels = c("Yes", "No")) +
  scale_x_discrete(labels = c("Atherosclerotic","Inflammatory"))
```


- Type of Aneurysm with Diameter: Inflammatory aneurysms are bigger.

```{r}
ggboxplot(subset(cases, !is.na(TIPOANEURISMA)), x = "TIPOANEURISMA", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(1,2), labels = c("Atherosclerotic","Inflammatory"))
```

- Symptoms with Smoking:

```{r}
tab <- as.data.frame(prop.table(table(cases$symptomatic, cases$Smoking)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Smoking Status", fill = "Symptoms") +
  scale_x_discrete(breaks = c("0","1"), labels = c("Asymptomatic","Symptomatic")) +
  scale_fill_discrete(breaks = c("0","1","2"),labels = c("Never", "Past", "Present"))
 
```


- Symptoms with Diameter: Symptomatic aneurysms are bigger.

```{r}
ggboxplot(subset(cases, !is.na(symptomatic)), x = "symptomatic", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(0,1), labels = c("Asymptomatic","Symptomatic"))
```


- Height with Weight.

```{r, warning=FALSE}
ggscatter(cases, x = "Weight_Kg", y = "Height_cm",
          add = "reg.line") +
  stat_cor(method = "pearson")
```


## Controls:

```{r}
controls <- bio2[98:nrow(bio2),]
controls <- controls[,2:ncol(controls)]
controls <- controls[,c("SEXO","age","Smoking")]
```

```{r, warning=FALSE}
M <- cor(controls, use = "pairwise.complete.obs")
testRes <- cor.mtest(controls, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

pdf(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correaltion_Plot_Controls.pdf")
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
dev.off()

corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
```


### Aneurysm type (Cases only):

```{r}
table(cases$TIPOANEURISMA, useNA = "always")
```

Among cases, there are 27 NAs

### Pathology (Controls only):

```{r}
#table(controls$Pathology, useNA = "always")
```

### Symptomatic (Cases only):

```{r}
table(cases$symptomatic, useNA = "always")
```

There are only 5 sympromatic aneurysms among cases.

Controls don't have information about Height or Weight.

### Height (Cases only):

```{r}
sum(is.na(cases$Height_cm))

17/nrow(cases)*100
```

```{r}
range(cases$Height_cm, na.rm = T)
```

```{r}
ggplot(cases, aes(Height_cm)) + geom_density()
```

There are 17 cases with no Height information.

### Weight (Cases only):

```{r}
sum(is.na(cases$Weight_Kg))
```

```{r}
range(cases$Weight_Kg, na.rm = T)
```

```{r}
ggplot(cases, aes(Weight_Kg)) + geom_density()
```

There are 16 cases with no Weight information.

### Smoking:

```{r}
table(cases$Smoking, useNA = "always")
```

There are 12 cases with no smoking information.

```{r}
table(controls$Smoking, useNA = "always")
```

There are 7 controls with no smoking information.

## Diseases

### Hypertension:

```{r}
table(bio$hypertension, bio$Status, useNA = "always")
```

There are 19 NAs.

### Dyslipedmia:

```{r}
table(bio$dyslipidemia, bio$Status, useNA = "always")
```

There are 19 NAs.

### Diabetes mellitus:

```{r}
table(bio$diabetes.mellitus, bio$Status, useNA = "always")
```

There are 19 NAs.

```{r}
if(identical(bio[is.na(bio$hypertension),]$Muestra, bio[is.na(bio$dyslipidemia),]$Muestra) == TRUE & identical(bio[is.na(bio$dyslipidemia),]$Muestra, bio[is.na(bio$diabetes.mellitus),]$Muestra) == TRUE){
  print("The 19 samples with NAs are the same for Hypertension, Dyslipedmia and Diabetes Mellitus")}
```

### PAD (only Cases):

```{r}
table(cases$PAD, useNA = "always")
```

There are 13 NAs among cases.

### Other Aneurysm (only cases):

```{r}
table(cases$other_aneurysm, useNA = "always")
```

There are 12 NAs among cases.

### CVE:

```{r}
table(bio$CVE, bio$Status, useNA = "always")
```

There are 21 NAs.

### CVD:

```{r}
table(bio$CVD, bio$Status, useNA = "always")
```

### COPD:

```{r}
table(bio$COPD, bio$Status, useNA = "always")
```

### Kidney Disease:

```{r}
table(bio$kidney_disease, bio$Status, useNA = "always")
```

### Aortic Diameter:

```{r}
ggplot(cases, aes(aortc_diameter_mm)) + geom_density()
```

There are 12 cases with NAs diameter.

```{r}
nrow(cases[cases$aortc_diameter_mm > 55,])
nrow(cases[cases$aortc_diameter_mm < 55,])
```

```{r}
range(cases$aortc_diameter_mm, na.rm = T)
```
















