---
title: "Study Correlations"
author: "Gerard Temprano Sagrera"
date: "17/01/2022"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo = FALSE}
library(openxlsx);library(ggplot2);library(corrplot);library(ggpubr);library(corrr);library(tidyverse)
```

```{r}
bio <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/20221115_Datos Pacientes.xlsx")
bio <- bio[2:nrow(bio),]
bio <- bio %>% remove_rownames %>% column_to_rownames(var="Muestra")
table(bio$Status)
```

```{r}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_CompleteV2.xlsx")
tec <- tec %>% remove_rownames %>% column_to_rownames(var="SampleID")
```

### Fix columns:

```{r}
# Select columns:
bio2 <- bio[,c("Status","SEXO","age","TIPOANEURISMA","symptomatic","Height_cm","Weight_Kg","Smoking","other_aneurysm",
                "aortc_diameter_mm")]

names(bio2) <- c("STATUS","SEX","AGE","TYPE","SYMPTOMS","HEIGHT","WEIGHT","SMOKING","O.ANEURYSM","DIAMATER")

# Transform to numeric:
bio2$STATUS <- as.numeric(as.factor(bio2$STATUS))
bio2$SEX <- as.numeric(bio2$SEX)
bio2$TYPE <- as.numeric(bio2$TYPE)
bio2$SYMPTOMS <- as.numeric(bio2$SYMPTOMS)
bio2$SMOKING <- as.numeric(bio2$SMOKING)
bio2$O.ANEURYSM <- as.numeric(bio2$O.ANEURYSM)
```


```{r}
tec2 <-  tec[ ,!names(tec) %in% c("Seq_ID","Case")]

names(tec2) <- c("FC","DATE","BATCH","GC MEAN","RIN","DV200","QUBIT")

# Transform to numeric:
tec2$FC <- as.numeric(as.factor(tec2$FC))
tec2$DATE <- as.numeric(as.factor(tec2$DATE))
```

```{r}
b.cases <- bio2[bio2$Status == "1",]
b.controls <- bio2[bio2$Status == "2",]

t.cases <- tec2[1:97,]
t.controls <- tec2[98:nrow(tec2),]
```


1.  Only sample, status, sex and age have no NAs.
2.  Cases only: Type, Symptomatic, Height, Weight, PAD, Other Aneurysm, Aortic Diameter, Leucocytes, Neutrophiles, Limphocytes, Monocytes, Platalets, Hb, RDW.
3.  Controls only: Pathology.


# Correlation study:

## ALL:

```{r}
# Remove pathology (only available in controls):
bio2 <- bio2[ ,!names(bio2) %in% c("Pathology")]
```

### Merge all covariates:

```{r}
all <- merge(bio2, tec2, by = 0)
all <- all %>% remove_rownames %>% column_to_rownames(var="Row.names")
```

```{r, warning=FALSE}
M <- cor(all, use = "pairwise.complete.obs")
testRes <- cor.mtest(all, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

#png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correlation_Plot_All.png", height = 10000, width = 10000, res = 1200)
#corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75)
#dev.off()

corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
```

- Type of aneurysm, height, weight, other aneurysm and size are only available for cases.

- Status with Sex: There are more male among cases.

```{r}
tab <- as.data.frame(prop.table(table(bio2$STATUS, bio2$SEX)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Status", fill = "Sex") +
  scale_x_discrete(breaks = c(1,2), labels = c("Case","Control")) +
  scale_fill_discrete(breaks = c(1,2), labels = c("Male","Female"))
```

- Status with Age: Cases are older.

```{r}
ggboxplot(bio2, x = "STATUS", y = "AGE") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 100) +
  scale_x_discrete(breaks = c(1,2), labels = c("Case","Control"))
```


- Status with Smoking: There are less smokers in controls.

```{r}
tab <- as.data.frame(prop.table(table(bio2$STATUS, bio2$SMOKING)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Status", fill = "Smoking") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Case","Control")) +
  scale_fill_discrete(breaks = c(0,1,2), labels = c("Never","Past","Currently"))
```


- Sex with Smoking: There are more smokers among men (There are more cases in men).

```{r}
tab <- as.data.frame(prop.table(table(bio2$SEX, bio2$SMOKING)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Sex", fill = "Smoking Status") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Man","Woman")) +
  scale_fill_discrete(breaks = c("0","1","2"), labels = c("Never","Past","Currently"))
```

- Age with Weight: Weight reduced with age.

```{r, warning=FALSE}
ggscatter(bio2, x = "AGE", y = "WEIGHT",
          add = "reg.line") +
  stat_cor(method = "pearson")
```

- Type of Aneurysm with Symptoms: There are more symptoms with Atherosclerotic aneurysms.

```{r}
tab <- as.data.frame(prop.table(table(bio2$TYPE, bio2$SYMPTOMS)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Type", fill = "Symptoms") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Atherosclerotic","Inflammatory")) +
  scale_fill_discrete(breaks = c("0","1"), labels = c("No", "Yes"))
```


- Type with Aortic Diameter: Inflammatory aneurysms have bigger diameters.

```{r}
ggboxplot(subset(bio2, !is.na(TYPE)), x = "TYPE", y = "DIAMATER") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(1,2), labels = c("Atherosclerotic","Inflammatory"))
```

- Symptoms with Smoking:

```{r}
tab <- as.data.frame(prop.table(table(bio2$SYMPTOMS, bio2$SMOKING)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Symptomatic", fill = "Smoking Status") +
  scale_x_discrete(breaks = c("0","1"), labels = c("Asymptomatic","Symptomatic")) +
  scale_fill_discrete(breaks = c("0","1","2"),labels = c("Never", "Past", "Present"))
```

- Symptoms with Diameter: Symptomatic aneurysms are bigger.

```{r}
ggboxplot(subset(bio2, !is.na(symptomatic)), x = "symptomatic", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(0,1), labels = c("Asymptomatic","Symptomatic"))
```

- Weight with Height.

```{r, warning=FALSE}
ggscatter(bio2, x = "Weight_Kg", y = "Height_cm",
          add = "reg.line") +
  stat_cor(method = "pearson")
```


## CASES:

### Merge all covariates:

```{r}
all.cases <- merge(b.cases, t.cases, by = 0)
all.cases <- all.cases %>% remove_rownames %>% column_to_rownames(var="Row.names")
```

```{r}
# Remove columns that have no variation:
all.cases <-  all.cases[ ,!names(all.cases) %in% c("Status","FC")]
```


```{r}
M <- cor(all.cases, use = "pairwise.complete.obs")
testRes <- cor.mtest(all.cases, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correlation_Plot_Cases.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
dev.off()

corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
```

```{r}
# Filter only biological samples:
filter.cases <-  all.cases[ ,c("SEXO","age","TIPOANEURISMA","symptomatic","Height_cm","Weight_Kg","Smoking","other_aneurysm","aortc_diameter_mm")]
```


```{r}
M <- cor(filter.cases, use = "pairwise.complete.obs")
testRes <- cor.mtest(filter.cases, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correlation_Plot_Cases_filter.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 2)
dev.off()

corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank")
```


- Age with Weight: Weight reduces with age.

```{r, warning=FALSE}
ggscatter(b.cases, x = "age", y = "Weight_Kg",
          add = "reg.line") +
  stat_cor(method = "pearson")
```


- Age with Smoking: Currently smokers are younger.

```{r}
my_comparisons <- list( c("0","1"), c("0", "2"), c("1", "2") )

ggboxplot(subset(b.cases, !is.na(Smoking)), x = "Smoking", y = "age", color = "Smoking") +
  stat_compare_means(comparisons = my_comparisons, label.y = c(90, 110, 130), method = "t.test") +
  stat_compare_means(label.y = c(150), method = "anova") +
  scale_x_discrete(breaks = c("0","1","2"), labels = c("Never","Past","Currently Smoking"))
```


- Type of Aneurysm with Symptoms: Atherosclerotic aneurysms have more Symptoms.

```{r}
tab <- as.data.frame(prop.table(table(b.cases$symptomatic, b.cases$TIPOANEURISMA)))

ggplot(tab, aes(x = Var2, y = Freq, fill=Var1)) + geom_col() + labs(fill = "Symptomatic", x = "Type") +
  scale_fill_discrete(labels = c("Yes", "No")) +
  scale_x_discrete(labels = c("Atherosclerotic","Inflammatory"))
```


- Type of Aneurysm with Diameter: Inflammatory aneurysms are bigger.

```{r}
ggboxplot(subset(b.cases, !is.na(TIPOANEURISMA)), x = "TIPOANEURISMA", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(1,2), labels = c("Atherosclerotic","Inflammatory"))
```

- Symptoms with Smoking:

```{r}
tab <- as.data.frame(prop.table(table(b.cases$symptomatic, b.cases$Smoking)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Smoking Status", fill = "Symptoms") +
  scale_x_discrete(breaks = c("0","1"), labels = c("Asymptomatic","Symptomatic")) +
  scale_fill_discrete(breaks = c("0","1","2"),labels = c("Never", "Past", "Present"))
 
```


- Symptoms with Diameter: Symptomatic aneurysms are bigger.

```{r}
ggboxplot(subset(b.cases, !is.na(symptomatic)), x = "symptomatic", y = "aortc_diameter_mm") +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 110) +
  scale_x_discrete(breaks = c(0,1), labels = c("Asymptomatic","Symptomatic"))
```


- Height with Weight.

```{r, warning=FALSE}
ggscatter(b.cases, x = "Weight_Kg", y = "Height_cm",
          add = "reg.line") +
  stat_cor(method = "pearson")
```


## CONTROLS:

### Merge all covariates:

```{r}
all.controls <- merge(b.controls, t.controls, by = 0)
all.controls <- all.controls %>% remove_rownames %>% column_to_rownames(var="Row.names")
```

```{r}
# Remove columns that have no variation:
all.controls <-  all.controls[ ,!names(all.controls) %in% c("Status","TIPOANEURISMA","symptomatic","Height_cm","Weight_Kg","other_aneurysm","aortc_diameter_mm")]
```

```{r, warning=FALSE}
M <- cor(all.controls, use = "pairwise.complete.obs")
testRes <- cor.mtest(all.controls, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

pdf(file = "C://Users/Gerard/Desktop/AAA/RNAseq/GitHub/Correlation_Plot_Controls.pdf")
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", addCoef.col = "black", number.cex = 0.6)
dev.off()

#corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", addCoef.col = "black", number.cex = 0.6)
```









