---
title: "Study Correlations"
author: "Gerard Temprano Sagrera"
date: "17/01/2022"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo = FALSE}
library(openxlsx);library(ggplot2);library(corrplot);library(ggpubr);library(corrr);library(tidyverse)
```

```{r}
bio <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/20221115_Datos Pacientes.xlsx")
bio <- bio[2:nrow(bio),]
bio <- bio %>% remove_rownames %>% column_to_rownames(var="Muestra")
table(bio$Status)
```

```{r}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_Complete.xlsx")
tec <- tec %>% remove_rownames %>% column_to_rownames(var="Seq_ID")
```

### Fix columns:

```{r}
# Select columns:
bio2 <- bio[,c("Status","SEXO","age","TIPOANEURISMA","symptomatic","Height_cm","Weight_Kg","Smoking","other_aneurysm",
                "aortc_diameter_mm")]

names(bio2) <- c("STATUS","SEX","AGE","TYPE","SYMPTOMS","HEIGHT","WEIGHT","SMOKING","O.ANEURYSM","DIAMETER")

# Transform to numeric:
bio2$STATUS <- as.numeric(as.factor(bio2$STATUS))
bio2$SEX <- as.numeric(bio2$SEX)
bio2$TYPE <- as.numeric(bio2$TYPE)
bio2$SYMPTOMS <- as.numeric(bio2$SYMPTOMS)
bio2$SMOKING <- as.numeric(bio2$SMOKING)
bio2$O.ANEURYSM <- as.numeric(bio2$O.ANEURYSM)
```


```{r}
tec2 <-  tec[ ,!names(tec) %in% c("SampleID")]

names(tec2) <- c("INDEX.i7","INDEX.i5","FC","LANE","LIBRARY","CASE","RIN","DV200","QUBIT","GC_MEAN")

# Transform to numeric:
tec2$INDEX.i5 <- as.numeric(as.factor(tec2$INDEX.i5))
tec2$INDEX.i7 <- as.numeric(as.factor(tec2$INDEX.i7))
tec2$FC <- as.numeric(as.factor(tec2$FC))
tec2$LANE <- as.numeric(as.factor(tec2$LANE))
```


1.  Only sample, status, sex and age have no NAs.
2.  Cases only: Type, Symptomatic, Height, Weight, PAD, Other Aneurysm, Aortic Diameter, Leucocytes, Neutrophiles, Limphocytes, Monocytes, Platalets, Hb, RDW.
3.  Controls only: Pathology.


# Correlation study:

## Cases vs Controls:

```{r}
# Remove pathology (only available in controls):
bio2 <- bio2[ ,!names(bio2) %in% c("Pathology")]
```

### Technical:

```{r}
M <- cor(tec2, use = "pairwise.complete.obs")
testRes <- cor.mtest(tec2, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Correlations/Correlation_Plot_All_Technical.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75, addCoef.col = "black")
dev.off()
```

### Biological:

```{r}
M <- cor(bio2, use = "pairwise.complete.obs")
testRes <- cor.mtest(tec2, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Correlations/Correlation_Plot_All_Biological.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75, addCoef.col = "black")
dev.off()
```


### Merge all covariates:

```{r}
all <- merge(bio2, tec2, by = 0)
all <- all %>% remove_rownames %>% column_to_rownames(var="Row.names")
```

```{r, warning=FALSE}
M <- cor(all, use = "pairwise.complete.obs")
testRes <- cor.mtest(all, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Correlations/Correlation_Plot_All.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75)
dev.off()
```


### Only Cases (1st Analysis):

```{r}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/First Analysis/Technical_CompleteV2.xlsx")
tec <- tec %>% remove_rownames %>% column_to_rownames(var="SampleID")

tec2 <-  tec[ ,!names(tec) %in% c("Seq_ID")]

names(tec2) <- c("FC","DATE","BATCH","CASE","GC_MEAN","RIN","DV200","QUBIT")

# Transform to numeric:
tec2$FC <- as.numeric(as.factor(tec2$FC))
tec2$DATE <- as.numeric(as.factor(tec2$DATE))
tec2$BATCH <- as.numeric(as.factor(tec2$BATCH))

tec2 <- tec2[tec2$CASE == 1,]

# All the same:
tec2 <-  tec2[ ,!names(tec2) %in% c("CASE", "FC")]
```

### Technical:

```{r}
M <- cor(tec2, use = "pairwise.complete.obs")
testRes <- cor.mtest(tec2, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Correlations/Correlation_Plot_Cases_Technical.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75, addCoef.col = "black")
dev.off()
```

### Biological:

```{r}
bio2 <- bio2[bio2$STATUS == "1",]

# All the same:
bio2 <-  bio2[ ,!names(bio2) %in% c("STATUS")]
```

```{r}
M <- cor(bio2, use = "pairwise.complete.obs")
testRes <- cor.mtest(bio2, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Correlations/Correlation_Plot_Cases_Biological.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75, addCoef.col = "black")
dev.off()
```

### Merge all covariates:

```{r}
all <- merge(bio2, tec2, by = 0)
all <- all %>% remove_rownames %>% column_to_rownames(var="Row.names")
```

```{r, warning=FALSE}
M <- cor(all, use = "pairwise.complete.obs")
testRes <- cor.mtest(all, conf.level = 0.95)

M[is.na(M)] <- 0
testRes$p[is.na(testRes$p)] <- 1

png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Correlations/Correlation_Plot_Cases.png", height = 10000, width = 10000, res = 1200)
corrplot(M, p.mat = testRes$p, sig.level = 0.05, type = "lower", insig = "blank", tl.cex = 1.75)
dev.off()
```



