---
title: "Data Statistical Description"
author: "Gerard Temprano Sagrera"
date: "2023-04-27"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo = FALSE}
library(openxlsx);library(ggplot2);library(corrplot);library(ggpubr);library(corrr);library(tidyverse)
```


```{r}
bio <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/20221115_Datos Pacientes.xlsx")
bio <- bio[2:nrow(bio),]
bio <- bio %>% remove_rownames %>% column_to_rownames(var="Muestra")

# Select columns:
bio2 <- bio[,c("Status","SEXO","age","TIPOANEURISMA","symptomatic","Height_cm","Weight_Kg","Smoking","other_aneurysm",
                "aortc_diameter_mm")]

names(bio2) <- c("STATUS","SEX","AGE","TYPE","SYMPTOMS","HEIGHT","WEIGHT","SMOKING","O.ANEURYSM","DIAMETER")

bio2[bio2$STATUS == "Case",]$STATUS <- "AAA"
```

- Type of aneurysm, height, weight, other aneurysm and size are only available for cases.

- Status with Sex: There are more male among cases.

```{r}
(tab <- as.data.frame(prop.table(table(bio2$STATUS, bio2$SEX), margin = 1)))
sum(!is.na(bio$SEXO))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "STATUS", fill = "Sex") +
   scale_fill_discrete(breaks = c(1,2), labels = c("Male","Female"))

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Status_Sex.png")
```

- Status with Age: Cases are older.

```{r}
sum(!is.na(bio2$STATUS))
sum(!is.na(bio2$AGE))

ggboxplot(bio2, x = "STATUS", y = "AGE") +
  geom_boxplot(aes(fill = factor(STATUS)), outlier.shape = " ") +
  geom_point(position = position_jitter(width = 0.2), color = "black", size = 2, alpha = 0.5) +
  stat_compare_means(method = "t.test", label.y = 100, label.x = 1.5) +
  theme(legend.position = "none")

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Status_Age.png")
```


- Status with Smoking: There are less smokers in controls.

```{r}
sum(!is.na(bio2$STATUS))
sum(!is.na(bio2$SMOKING))

tab <- as.data.frame(prop.table(table(bio2$STATUS, bio2$SMOKING), margin = 1))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "STATUS", fill = "Smoking") +
  scale_fill_discrete(breaks = c(0,1,2), labels = c("Never","Past","Currently"))

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Status_Smoking.png")
```


```{r}
bio.cases <- bio2[bio2$STATUS == "Case",]
```



- Type of Aneurysm with Symptoms: There are more symptoms with Inflammatory aneurysms.

```{r}
sum(!is.na(bio.cases$TYPE) & !is.na(bio.cases$SYMPTOMS))

tab <- as.data.frame(prop.table(table(bio.cases$TYPE, bio.cases$SYMPTOMS), margin = 1))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Type", fill = "Symptoms") +
  scale_x_discrete(breaks = c("1","2"), labels = c("Atherosclerotic","Inflammatory")) +
  scale_fill_discrete(breaks = c("0","1"), labels = c("No", "Yes"))

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Symptoms_Type.png")
```


- Type with Aortic Diameter: Inflammatory aneurysms have bigger diameters.

```{r}
sum(!is.na(bio.cases$TYPE) & !is.na(bio.cases$DIAMETER))
table(bio.cases$TYPE)

tab <- as.data.frame(prop.table(table(bio.cases$TYPE, bio.cases$DIAMETER), margin = 1))

ggboxplot(subset(bio.cases, !is.na(TYPE)), x = "TYPE", y = "DIAMETER", outlier.shape = NA) +
  geom_boxplot(aes(fill = factor(TYPE)), outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.2), color = "black", size = 2, alpha = 0.5) +
  stat_compare_means(method = "t.test", label.y = 110, label.x = 1.4) +
  scale_x_discrete(breaks = c(1, 2), labels = c("Atherosclerotic", "Inflammatory")) +
  theme(legend.position = "none")

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Type_Diameter.png")
```


- Symptoms with Diameter: Symptomatic aneurysms are bigger.

```{r}
sum(!is.na(bio.cases$SYMPTOMS) & !is.na(bio.cases$DIAMETER))

ggboxplot(subset(bio.cases, !is.na(SYMPTOMS)), x = "SYMPTOMS", y = "DIAMETER", outlier.shape = NA) +
geom_boxplot(aes(fill = SYMPTOMS), outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.2), color = "black", size = 2, alpha = 0.5) +
  stat_compare_means(method = "t.test", label.y = 110, label.x = 1.3) +
  scale_x_discrete(breaks = c(0,1), labels = c("Asymptomatic","Symptomatic")) +
  theme(legend.position = "none")

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Symptoms_Diameter.png")
```



- Symptoms with Smoking:

```{r}
tab <- as.data.frame(prop.table(table(bio.cases$SYMPTOMS, bio.cases$SMOKING)))

ggplot(tab, aes(x = Var1, y = Freq, fill=Var2)) + geom_col() + labs(x = "Smoking Status", fill = "Symptoms") +
  scale_x_discrete(breaks = c("0","1"), labels = c("Asymptomatic","Symptomatic")) +
  scale_fill_discrete(breaks = c("0","1","2"),labels = c("Never", "Past", "Present"))

ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Symptoms_Smoking.png")
```


- Age with Type:

```{r}
sum(!is.na(bio.cases$AGE) & !is.na(bio.cases$TYPE))

ggboxplot(subset(bio.cases, !is.na(TYPE)), x = "TYPE", y = "AGE") +
geom_boxplot(aes(fill = TYPE)) +
  geom_point(position = position_jitter(width=0.2), color = "black", size = 2, alpha = 0.5) +
  stat_compare_means(method = "t.test", label.y = 110, label.x = 1.3) +
  scale_x_discrete(breaks = c(1,2), labels = c("Inflammatory","Atherosclerotic")) +
  theme(legend.position = "none")

#ggsave("C://Users/Gerard/Desktop/AAA/RNAseq/Plots/SuppFig3/Age_Type.png")
```


- Weight with Type:

```{r}
sum(!is.na(bio.cases$WEIGHT) & !is.na(bio.cases$TYPE))

ggboxplot(subset(bio.cases, !is.na(TYPE) & !is.na(WEIGHT)), x = "TYPE", y = "WEIGHT") +
geom_boxplot(aes(fill = TYPE)) +
    geom_point(position = position_jitter(width=0.2), color = "black", size = 2, alpha = 0.5) +
  stat_compare_means(method = "t.test", label.y = 110, label.x = 1.3) +
  scale_x_discrete(breaks = c(1,2), labels = c("Inflammatory","Atherosclerotic")) +
  theme(legend.position = "none")
```



