---
title: "RNAseq AAA Normalization Cases"
author: "Gerard Temprano Sagrera"
date: "2022-12-16"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(edgeR);library(openxlsx);library(data.table);library(DESeq2);library(tximport);library(stringr);library(ggbiplot);library(dplyr);library(knitr);library(gridExtra);library(corrplot);library(tidyverse)
```

### Apply TPM transformation to salmon counts:

```{r}
# Load previously prepared salmon read counts:
counts <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Counts.salmon.txt")
lengths <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Lengths.txt")
lengths <- lengths$length/1000
rnames <- counts$V1
counts <- counts[,-1]
rownames(counts) <- rnames

r_tpm <- function(dfr,len)
{
  dfr1 <- sweep(dfr,1,len,`/`)
  sc.f <- colSums(dfr1)/(10^6)
  return(sweep(dfr1,2,sc.f,`/`))
}

counts.tpm <- as.data.frame(r_tpm(counts, lengths))
```

### Load Technical table and save a case/control variable:

```{r}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_CompleteV2.xlsx")
cc <- tec[,c("Case")]
cc[cc == 1] <- "Case"
cc[cc == 0] <- "Control"
```

### Filter lowely expressed genes:

Keep genes that have more than 50% of non-zero counts across all samples:
```{r}
keep <- apply(counts.tpm, 1, function(x) sum(x == 0) <= 70)
sum(keep) # 21,355

counts.tpm <- counts.tpm[keep,]

# Check how this filter affects cases and controls:
cases.tpm <- counts.tpm[,1:97]
controls.tpm <- counts.tpm[,98:ncol(counts.tpm)]

sum(apply(cases.tpm, 1, function(x) sum(x == 0) <= 48)) # 21,197
sum(apply(controls.tpm, 1, function(x) sum(x == 0) <= 22)) # 20,499
```

### Keep genes that have more than 50% of non-zero counts across cases and across controls:

```{r}
#cases.counts <- counts.tpm[,1:97]
#controls.counts <- counts.tpm[,98:ncol(counts.tpm)]

#keep.cases <- apply(cases.counts, 1, function(x) sum(x == 0) <= 48)
#sum(keep.cases) # 21,582

#keep.controls <- apply(controls.counts, 1, function(x) sum(x == 0) <= 22)
#sum(keep.controls) # 20,922

#length(intersect(rownames(cases.counts[keep.cases, ]), rownames(controls.counts[keep.controls, ]))) # 20,341
```

### Transform to DGEList format:

```{r}
dge <- DGEList(counts=cases.tpm)
```

### Apply TMM Normalization:

```{r}
dge <- calcNormFactors(dge, method = "TMM")
```

### Apply logCPM transformation to test the effects of covariates:

```{r}
dgeTpmAdjCpm <- cpm(dge)
```

### Perform PCA:

```{r}
my_pca <- prcomp(t(dgeTpmAdjCpm), center = T, scale. = T)
summaryPca <- summary(my_pca)
```

### PCA Plots:

```{r, include=TRUE}
barplot((my_pca$sdev^2 / sum(my_pca$sdev^2))[1:20], names.arg = 1:20,  
        xlab = "PC #", ylab = "Amount of explained variance per PC",
        ylim=c(0,0.4), las=2, cex.names=.8)
```

```{r, include=TRUE}
plot(cumsum(my_pca$sdev^2 / sum(my_pca$sdev^2))[1:10], type="b", ylab = "Cummulative proportion of Variance Explained")
```

### PCA Plot Controls:

```{r, include=TRUE}
ggbiplot(my_pca, labels = colnames(dgeTpmAdjCpm), var.axes = F, obs.scale = 1, var.scale = 1) +
  scale_color_discrete(name = "") +
  theme(legend.direction = "horizontal", legend.position = "top")
```

### Plot the histograms of the 9 smallest PC1:

```{r}
#par(mfrow = c(3, 3))
#pca1.min <- head(order(predict(my_pca)[,1]), 9)
#for (i in 1:9) hist(as.numeric(dgeTpmAdjCpm[pca1.min[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

### Plot the histograms of the 9 largest PC1:

```{r}
#par(mfrow = c(3, 3))
#pca1.max <- head(order(predict(my_pca)[,1], decreasing = T), 9)
#for (i in 1:9) hist(as.numeric(dgeTpmAdjCpm[pca1.max[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

### Technical batches names:

```{r, echo = FALSE}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_CompleteV2.xlsx")

# Select only cases:
tec <- tec[1:97,]
```

### Covariates against PCA:

#### Date:

```{r}
covariate <- "Date"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- c()
Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Batch:

```{r}
covariate <- "Batch"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### GC_Mean:

```{r}
covariate <- "GC_Mean"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### RIN:

```{r}
covariate <- "RIN"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Prepare and save Correlations table:

```{r}
rownames(Correlations) <- paste0("PC",1:10)
colnames(Correlations) <- c("Date","Batch","GC_Mean","RIN")

write.xlsx(as.data.frame(Correlations), "C://Users/Gerard/Desktop/AAA/RNAseq/Correlation_Cases.xlsx", sep = "\t", rowNames = T)
```

Recommended technical variables: Date, Batch, GC_Mean.

```{r}
tec2 <- tec %>% remove_rownames %>% column_to_rownames(var="SampleID")
tec2 <- tec2[,2:ncol(tec2)]

tec2$Date <- as.numeric(as.factor(tec2$Date))
tec2 <- tec2[,c("Date","Batch","GC_Mean","RIN","DV200","Qubit")]


corrplot(cor(tec2), method = "square", type = "lower", order = "AOE", addCoef.col = "black", tl.pos = "ld", cl.pos = "n")
```

### GC Content Histogram:

```{r}
hist(tec$GC_Mean, 
     breaks = 40, 
     main = "GC_mean histogram", 
     xlab = "GC_mean")
```

### Extreme Gc Content values:

```{r}
# Left tail
par(mfrow = c(3, 3))
gc.min <- head(order(tec$GC_Mean), 9)
for (i in 1:9) hist(as.numeric(dgeTpmAdjCpm[gc.min[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

```{r}
# Right tail
par(mfrow = c(3, 3))
gc.max <- head(order(tec$GC_Mean, decreasing = T), 9)
for (i in 1:9) hist(as.numeric(dgeTpmAdjCpm[gc.max[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

### Covariates influence in the PCA:

```{r}
df.predict <- data.frame(predict(my_pca))[, 1:2]
df.predict$SampleID <- rownames(df.predict)
df.all <- plyr::join(tec, df.predict, by = "SampleID", type = "right")
```

### Batch = Plate:

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(Batch)), size = 2.5) + 
  ggtitle("Technical variable: Plate") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(colour = "Plate")
```

### Date:

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(Date)), pch = 20, size = 3) + 
  ggtitle("Technical variable: Date", )+
  theme(plot.title = element_text(hjust = 0.5))
```

### GC Content:

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(GC_Mean), shape = as.factor(Case)), size = 2.5) +
  ggtitle("Technical variable: GC Content", )+
  theme(plot.title = element_text(hjust = 0.5))
```

### RIN:

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = RIN), pch = 20, size = 3) + 
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Technical variable: RIN")
```

### DV200:

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = DV200), pch = 20, size = 3) + 
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  ggtitle("Technical variable: DV200")
```

### Qubit:

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = Qubit), pch = 20, size = 3) + 
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  ggtitle("Technical variable: Quvit")
```
