---
title: "RNAseq AAA Normalization"
author: "Gerard Temprano Sagrera"
date: "07/12/2022"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(edgeR);library(openxlsx);library(data.table);library(DESeq2);library(tximport);library(stringr);library(ggbiplot);library(dplyr);library(knitr);library(gridExtra)
```

### Apply TPM transformation to salmon counts:

```{r}
# Load previously prepared salmon read counts:
counts <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Counts.salmon.txt")
lengths <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Lengths.txt")
lengths <- lengths$length/1000
rnames <- counts$V1
counts <- counts[,-1]
rownames(counts) <- rnames

r_tpm <- function(dfr,len)
{
  dfr1 <- sweep(dfr,1,len,`/`)
  sc.f <- colSums(dfr1)/(10^6)
  return(sweep(dfr1,2,sc.f,`/`))
}

counts.tpm <- as.data.frame(r_tpm(counts, lengths))
```

### Use EdgeR to normalize the counts:

```{r, echo = FALSE}
#cc <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Techinc_1.xlsx")
#group <- cc[,c("Case/Control")]

# Check they have the same order:
#identical(colnames(counts),cc$Nombre.de.la.muestra)

#y <- DGEList(counts, group = group)
#dim(y)

#y2 <- calcNormFactors(y)
```

### Use deseq2 to normalize the counts:

```{r, echo = FALSE}
#cc <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Techinc_1.xlsx")
#group <- cc[,c("Nombre.de.la.muestra","Case/Control")]
#colnames(group) <- c("ID","Status")
#rownames(group) <- group$ID
#group$Status <- factor(group$Status, levels = c("1","0"), labels = c("Case","Control"))
#group$Status <- relevel(group$Status, ref = "Control")

# Check they have the same order:
#identical(rownames(group),colnames(counts))
#all(rownames(group$ID) == colnames(counts))

#y <- DESeqDataSetFromMatrix(counts, colData = group, design = ~ Status)
#dim(y)

#keep <- rowSums(counts(y)) >= 10
#y <- y[keep,]
#counts(y)
```

### Perform PCA, removing genes that do not have variability:

```{r}
counts.tpm <- as.data.frame(r_tpm(counts, lengths))
```

```{r}
tec2 <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_CompleteV2.xlsx")
cc <- tec2[,c("Case")]
cc[cc == 1] <- "Case"
cc[cc == 0] <- "Control"
counts.tpm <- t(counts.tpm)

# Remove genes that have 0 variability:
counts_var <- counts.tpm[,apply(counts.tpm, 2, var, na.rm=TRUE) != 0]

my_pca <- prcomp(counts_var, center = T, scale. = T)
```

### PCA Plots:

```{r, include=TRUE}
plot(my_pca, main = "Variance Explained")
```

```{r, include=TRUE}
plot(cumsum(my_pca$sdev^2 / sum(my_pca$sdev^2))[1:10], type="b", ylab = "Cummulative proportion of Variance Explained")
```

### PCA Plot:

```{r, include=TRUE}
ggbiplot(my_pca, labels = rownames(counts_var), groups = cc, var.axes = F, obs.scale = 1, var.scale = 1) +
  scale_color_discrete(name = "") +
  theme(legend.direction = "horizontal", legend.position = "top")
```

### Plot the histograms of the 9 smallest PC1:

```{r}
par(mfrow = c(3, 3))
pca1.min <- head(order(predict(my_pca)[, 1]), 9)
for (i in 1:9) hist(as.numeric(counts_var[pca1.min[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

### Plot the histograms of the 9 largest PC1:

```{r}
par(mfrow = c(3, 3))
pca1.max <- head(order(predict(my_pca)[, 1], decreasing = T), 9)
for (i in 1:9) hist(as.numeric(counts_var[pca1.max[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

### Technical batches names:

```{r, echo = FALSE}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_CompleteV2.xlsx")
```

### Covariates against PCA:

#### Case

```{r}
covariate <- "Case"
boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) cor.test(boxplotMatrix[,x],as.numeric(boxplotMatrix[,covariate]), method="spearman", exact = FALSE)[c(3,4)])), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "correlation")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$correlation <- as.numeric(format(dfPvalue$correlation, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  xlab(covariate) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- c()
Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Date:

```{r}
covariate <- "Date"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### FC:

```{r}
covariate <- "FC"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Batch:

```{r}
covariate <- "Batch"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### GC_Mean:

```{r}
covariate <- "GC_Mean"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### RIN:

```{r}
covariate <- "RIN"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrixMelt, aes(x=as.factor(covariate), y=value)) +
  geom_boxplot() +
  theme_bw(base_size = 16) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~variable,  ncol=2)
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Prepare and save Correlations table:

```{r}
rownames(Correlations) <- paste0("PC",1:10)
colnames(Correlations) <- c("Case","Date","FC","Batch","GC_Mean","RIN")

write.xlsx(as.data.frame(Correlations), "C://Users/Gerard/Desktop/AAA/RNAseq/Correlation.xlsx", sep = "\t", rowNames = T)
```

```{r}
table(tec$Batch, tec$Date)
```

```{r}
table(tec$Batch, tec$FC)
```

```{r}
table(tec$FC, tec$Date)
```

GC Content Histogram

```{r}
hist(tec$GC_Mean, 
     breaks = 40, 
     main = "GC_mean histogram", 
     xlab = "GC_mean")
```

Extreme Gc Content values

```{r}
# Left tail
par(mfrow = c(3, 3))
gc.min <- head(order(tec$GC_Mean), 9)
for (i in 1:9) hist(as.numeric(counts_var[gc.min[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

```{r}
# Right tail
par(mfrow = c(3, 3))
gc.max <- head(order(tec$GC_Mean, decreasing = T), 9)
for (i in 1:9) hist(as.numeric(counts_var[gc.max[i], ]), main = paste0("Number ", i), xlab = "Gene expression")
```

# Covariates influence in the PCA:

```{r}
df.predict <- data.frame(predict(my_pca))[, 1:2]
df.predict$SampleID <- rownames(df.predict)
df.all <- plyr::join(tec, df.predict, by = "SampleID", type = "right")
```

### Batch = Plate

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(Batch), shape = as.factor(Case)), size = 2.5) +
  ggtitle("Technical variable: Batch") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Batch = Plate

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(Batch)), pch = 20, size = 3) + 
  ggtitle("Technical variable: Plate") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(colour = "Plate")
```

### Flow Cell

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = FC, shape = as.factor(Case)), size = 2.5) +
  ggtitle("Technical variable: Flow Cell", )+
  theme(plot.title = element_text(hjust = 0.5))
```

### Flow Cell

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = FC), pch = 20, size = 3) + 
  ggtitle("Technical variable: Flow Cell", )+
  theme(plot.title = element_text(hjust = 0.5))
```

### Date

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(Date)), pch = 20, size = 3) + 
  ggtitle("Technical variable: Date", )+
  theme(plot.title = element_text(hjust = 0.5))
```

### GC Content

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = as.factor(GC_Mean), shape = as.factor(Case)), size = 2.5) +
  ggtitle("Technical variable: GC Content", )+
  theme(plot.title = element_text(hjust = 0.5))
```

### RIN

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = RIN), pch = 20, size = 3) + 
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Technical variable: RIN")
```

### DV200

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = DV200), pch = 20, size = 3) + 
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  ggtitle("Technical variable: DV200")
```

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = DV200, shape = as.factor(Case)), size = 2.5) +
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  ggtitle("Technical variable: DV200")
```

### Qubit

```{r}
ggplot(df.all, aes(x = PC1, y = PC2)) + 
  theme_bw() + 
  coord_fixed() + 
  geom_point(aes(colour = Qubit), pch = 20, size = 3) + 
  scale_colour_continuous(low = "lightblue", high = "darkblue") + 
  ggtitle("Technical variable: Quvit")
```
