---
title: "1.3.Manual_Analysis_RSEM_Cases"
author: "Gerard Temprano Sagrera"
date: "2023-02-24"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(data.table);library(edgeR);library(tidyverse);library(openxlsx);library(preprocessCore);library("org.Hs.eg.db");library(stats);library(qvalue);library(ggpubr);library(VennDiagram);library(grid)
```

### Load Annotation file to change format names:

```{r}
ann <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Annotation/gencode.v26.annotation.fixed.gtf.gz")

egfr.id <- ann[ann$gene_name %in% "EGFR",]$gene_id
```

### Counts are already in TPM from RSEM:

```{r}
counts.tpm <- read.table("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Counts.RSEM.Cases.txt")
```

### Subset significant genes when comparing cases vs controls:

```{r}
#sig.genes <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Significant_genes.txt", header = F)
#sig.genes <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Significant_genes_smoking.txt", header = F)

#counts.tpm <- counts.tpm[rownames(counts.tpm) %in% sig.genes$V1,]
```

### Load technical data and subset to cases:

```{r}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/First Analysis/Technical_CompleteV2.xlsx")
tec <- tec[tec$SampleID %in% colnames(counts.tpm),]
```

### Load biological data and subset to cases:

```{r}
bio <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/20221115_Datos Pacientes.xlsx")
bio <- bio[bio$Muestra %in% colnames(counts.tpm),]

# Smoking:
bio$Smoking2 <- ifelse(bio$Smoking == 0, 0, 1)
```


### By Diameter:

```{r}
dge <- DGEList(counts = counts.tpm)

egfr <- t(dge$counts)[,egfr.id]
```

``


```{r}
design <-
  model.matrix(
    ~ egfr + tec$Date + tec$Batch + tec$GC_Mean + tec$RIN + tec$DV200 + tec$Qubit + bio$age)

colnames(design) <-
  c("(Intercept)",
    "EGFR",
    "Date",
    "Batch",
    "GC_Mean",
    "RIN",
    "DV200",
    "Qubit",
    "Age")
```

### Remove lowely expressed genes using LIMMA:

```{r}
keep <- filterByExpr(dge, design)

paste0("We keep ", sum(keep), " genes") # 10,575.
paste0("We remove ", sum(!keep), " genes") # 16,715.

dge <- dge[keep,,keep.lib.sizes=FALSE]
```

### Apply quantile normalization to counts:

```{r}
counts.tpm <- as.data.frame(normalize.quantiles(as.matrix(dge$counts), keep.names = TRUE))
```

### Manual Differential Expression:

```{r}
print(paste0("There are ",sum(is.na(egfr))," NAs"))

Pvalue <-
  apply(counts.tpm, 1, function(x)
    summary(
      lm(
        x ~ egfr + tec$Date + tec$Batch + tec$GC_Mean + tec$RIN + tec$DV200 + tec$Qubit + bio$age
      )
    )[4][[1]][2, 4])
Coef <-
  apply(counts.tpm, 1, function(x)
    summary(
      lm(
        x ~ egfr + tec$Date + tec$Batch + tec$GC_Mean + tec$RIN + tec$DV200 + tec$Qubit + bio$age
      )
    )[4][[1]][2, 1])
```


### Plot P-values:

```{r}
#png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Hist_Diameter.png", height = 10000, width = 10000, res = 1200)
#hist(Pvalue, main = "", cex.main = 2)
#dev.off()

hist(Pvalue, main = "EGFR Cases", cex.main = 2)
```

### Create qvalue object:

```{r}
qobj <- qvalue(p = Pvalue)
```

```{r, warning=FALSE}
#summary(qobj)
#hist(qobj)
#plot(qobj)
```

```{r}
Results <- data.frame(
  pvalue = Pvalue,
  qvalue = qobj$qvalues,
  fdr = p.adjust(Pvalue, method = "fdr"),
  pi1 = 1-qobj$pi0,
  Beta = Coef
)

Results <- merge(Results, ann[,c("gene_id","gene_name")], by.x = 0, by.y = "gene_id")
names(Results)[names(Results) == "Row.names"] <- "gene_id"
Results <- Results[!duplicated(Results$gene_name),]
Results <- Results %>% remove_rownames %>% column_to_rownames(var="gene_name")
Results <- Results[order(Results$pvalue),]

nrow(Results[Results$fdr < 0.05,]) # 4219

#fwrite(Results[Results$fdr < 0.05,], "C://Users/Gerard/Desktop/AAA/RNAseq/SigRes/Diameter.txt", sep = "\t", row.names = T)
```


### Functional enrichment analysis:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(clusterProfiler);library(org.Hs.eg.db);library(enrichplot)
```

```{r}
gene_list <- substr(Results[Results$fdr < 0.05,]$gene_id, 1, 15)

entrez_ids <- bitr(gene_list, fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# GO enrichment analysis:
go_enrich <- enrichGO(gene          = entrez_ids$ENTREZID,
                      OrgDb         = org.Hs.eg.db,
                      keyType       = "ENTREZID",
                      ont           = "ALL", # Choose from "BP" (Biological Process), "CC" (Cellular Component), or "MF" (Molecular Function)
                      pAdjustMethod = "BH",
                      pvalueCutoff  = 0.05,
                      qvalueCutoff  = 0.2)
```

```{r}
# GO enrichment dot plot:
png("C://Users/Gerard/Desktop/GO_Enrichment_EGFR.png", width = 4*800, height = 4*600, res = 300)
dotplot(go_enrich, showCategory = 20) + ggtitle("GO Enrichment")
dev.off()

go_enrich@result
```

```{r}
gene_id_list <- lapply(go_enrich@result$geneID, function(x) strsplit(x, "/")[[1]])

gene_symbols_df <- bitr(unlist(gene_id_list), fromType = "ENTREZID", toType = "SYMBOL", OrgDb = org.Hs.eg.db)

gene_symbol_list <- lapply(gene_id_list, function(x) gene_symbols_df$SYMBOL[match(x, gene_symbols_df$ENTREZID)])
names(gene_symbol_list) <- names(gene_id_list)

collapsed_gene_symbols <- sapply(gene_symbol_list, function(x) paste(x, collapse = ","))

res <- cbind(go_enrich@result, collapsed_gene_symbols)

res[grep("EGFR",res$collapsed_gene_symbols),]

res[res$collapsed_gene_symbols]
```




```{r}
gene_list <- rownames(Results[Results$fdr < 0.05,])

# KEGG enrichment analysis
kegg_enrich <- enrichKEGG(gene         = entrez_ids$ENTREZID,
                          organism     = "hsa",
                          keyType      = "kegg",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.2)
```

```{r}
# KEGG enrichment dot plot
png("C://Users/Gerard/Desktop/KEGG_Enrichment_EGFR.png", width = 4*800, height = 4*600, res = 300)
dotplot(kegg_enrich, showCategory = 15) + ggtitle("KEGG Enrichment")
dev.off()
```


#### EGFR:

```{r}
counts <- merge(counts.tpm, ann[,c("gene_id","gene_name")], by.x = 0, by.y = "gene_id")
counts <- counts[,-1]
counts <- counts[!duplicated(counts$gene_name),]
counts <- counts %>% remove_rownames %>% column_to_rownames(var="gene_name")

counts <- as.matrix(counts)
```

```{r}
res1 <- resid(lm(counts["EGFR",] ~ tec$Date + tec$Batch + tec$GC_Mean + tec$RIN + tec$DV200 + tec$Qubit + bio$age))
```



# For loop for multiple genes:
```{r}
genes <- rownames(Results)[1:10]

i=1

for (i in 1:length(genes)){
  res2 <- resid(lm(counts[genes[i],] ~ tec2$Date + tec$Batch + tec$GC_Mean + tec$RIN + tec$DV200 + tec$Qubit + bio$age))
  res <- data.frame(EGFR = res1,
                  Gene2 = res2)
  res$row_names <- rownames(res)

  p <- ggplot(res, aes(x = EGFR, y = Gene2)) + 
  geom_text(aes(label = row_names), color = "orangered2", check_overlap = TRUE) + 
  ggtitle(genes[i]) + 
  ylab (genes[i]) + 
  xlab("EGFR") +
  geom_smooth(method = 'lm', formula = y ~ x, color = "black") + 
  stat_cor(method = "pearson") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 22),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 20),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )
print(p)
}
```



# For one gene:

```{r}
res2 <- resid(lm(counts["NUMBL",] ~ tec$Date + tec$Batch + tec$GC_Mean + tec$RIN + tec$DV200 + tec$Qubit + bio$age))

res <- data.frame(EGFR = res1,
                  NUMBL = res2)

res$row_names <- rownames(res)

ggplot(res, aes(x = EGFR, y = NUMBL)) + 
  geom_text(aes(label = row_names), color = "orangered2", check_overlap = TRUE) + 
  ggtitle(genes[i]) + 
  ylab(genes[i]) + 
  xlab("EGFR") +
  geom_smooth(method = 'lm', formula = y ~ x, color = "black") + 
  stat_cor(method = "pearson") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 22),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 20),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )

```


















