---
title: "1.1.Sex_Identification"
author: "Gerard Temprano Sagrera"
date: "2023-02-27"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(data.table);library(stringr);library(openxlsx);library(stringr);library(ggbiplot);library(dplyr);library(knitr);library(gridExtra);library(tidyverse)
```

### Counts are already in TPM from RSEM:

```{r}
genes <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Annotation/gencode.v26.annotation.fixed.gtf.gz")

counts.tpm <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Counts.RSEM.txt")
counts.tpm <- counts.tpm %>% remove_rownames %>% column_to_rownames(var="V1")
```


```{r}
# I am using the raw counts to increase the number of available genes:
xist <- genes[genes$gene_name=="XIST",]$gene_id
counts.sex.f <- counts.tpm[rownames(counts.tpm)%in%xist,]

y.gene <- c("SRY","ZFY","RPS4Y1","AMELY","TBL1Y","PCDH11Y","TGIF2LY","TSPY1","TSPY2","AZFa","USP9Y","DDX3Y","UTY","TB4Y","AZFb","CYorf15","RPS4Y2","EIF1AY","KDM5D","XKRY","PRY","PRY2","HSFY1","HSFY2","RBMY1A1","AZFc","DAZ1","DAZ2","DAZ3","DAZ4","CDY1","CDY2","VCY1","VCY2")

y.list <- genes[genes$gene_name%in%y.gene,]$gene_id
counts.sex.m <- counts.tpm[rownames(counts.tpm)%in%y.list,]

# Combine the XIST gene with the mean expression of all Y chromosome genes:
counts.sex <- rbind(counts.sex.f, data.frame(t(colMeans(counts.sex.m))))
counts.sex <- as.data.frame(t(counts.sex))
colnames(counts.sex) <- c("XIST","Chromosome_Y")
```


### Apply log transformation to the sex counts:

```{r}
counts.sex.log <- log(counts.sex)
counts.sex.log[sapply(counts.sex.log, is.infinite)] <- 0
```

### Apply normalization to the sex counts:

```{r}
counts.sex.log.scaled <- as.data.frame(scale(counts.sex.log))
colMeans(counts.sex.log.scaled)
apply(counts.sex.log.scaled, 2, sd)
counts.sex.log.scaled$ID <- rownames(counts.sex.log.scaled)
```

```{r}
sex <- read.xlsx("C://Users/Gerard/Downloads/20221115_Datos Pacientes.xlsx")
sex <- sex[2:nrow(sex),]
counts.sex.log.scaled <- merge(counts.sex.log.scaled, sex[,c("Muestra","SEXO")], by.x = "ID", by.y = "Muestra")
names(counts.sex.log.scaled)[names(counts.sex.log.scaled) == "SEXO"] <- "Pheno.Sex"
```


```{r, fig.show="hold", out.width="50%"}
counts.sex.log.scaled$label_row <- ifelse(counts.sex.log.scaled$ID == "HSP075", "yes", "no")

Male <- ggplot(data = counts.sex.log.scaled, aes(x=XIST, y=Chromosome_Y, color=Pheno.Sex, label = ifelse(label_row == "yes", ID, ""))) +
  theme_bw() +
  ggtitle("Gender RNAseq") +
  scale_color_manual(labels = c("Male", "Female"), values = c("lightblue", "red")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point() +
  geom_text(data = subset(counts.sex.log.scaled, label_row == "yes"), size = 3, show.legend = FALSE)

Female <- ggplot(data = counts.sex.log.scaled, aes(x=XIST, y=Chromosome_Y, color=Pheno.Sex, label = ifelse(label_row == "yes", ID, "")))  +
  theme_bw() +
  ggtitle("Gender RNAseq") +
  scale_color_manual(labels=c("Male", "Female"), values=c("red", "lightblue")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point() +
  geom_text(data = subset(counts.sex.log.scaled, label_row == "yes"), size = 3, show.legend = FALSE)
```


```{r}
ggsave(file="C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Sex_Identification1.png", Male)
ggsave(file="C://Users/Gerard/Desktop/AAA/RNAseq/Plots/Sex_Identification2.png", Female)
```

### Check the sex between the controls we could not identify:

```{r}
counts.sex[rownames(counts.sex) %in% c("HSP214", "HSP217"),]
```
We determine that HSP214 is the Male and HSP217 is the Female.

### Check the values of the mismatched sample:

```{r}
counts.sex[rownames(counts.sex) %in% "HSP075",]
```
It clearly looks like a Male, although it is classified as a Female in the biological information.

### Create column of sex determined by the expression (Seq.Sex):

```{r}
counts.sex.log.scaled$Seq.Sex <- ifelse(counts.sex.log.scaled$XIST > 1 | counts.sex.log.scaled$Chromosome_Y < -0.5, 2, 1 )
```

### Mismatched samples:

```{r}
counts.sex.log.scaled[!counts.sex.log.scaled$Pheno.Sex == counts.sex.log.scaled$Seq.Sex,]
```


















