---
title: "1.2.Normalization_PCA_Covariates.Rmd"
author: "Gerard Temprano Sagrera"
date: "2023-02-23"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(edgeR);library(openxlsx);library(data.table);library(DESeq2);library(tximport);library(stringr);library(ggbiplot);library(dplyr);library(knitr);library(gridExtra);library(corrplot);library(tidyverse)
```


### Counts are already in TPM from RSEM:

```{r}
counts.tpm <- read.table("C://Users/Gerard/Desktop/AAA/RNAseq/Gene_counts/Counts.RSEM.All.txt")
```


### Load Technical table and save a case/control variable:

```{r}
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_Complete.xlsx")
tec <- tec[tec$Seq_ID %in% colnames(counts.tpm),]
cc <- tec[,c("Case")]
cc[cc == 1] <- "Case"
cc[cc == 0] <- "Control"

# Transform to factor:
tec$Case <- as.factor(tec$Case)
tec$Library <- as.factor(tec$Library)

# Transform to numeric:
tec$index.i7 <- as.numeric(as.factor(tec$index.i7))
tec$index.i5 <- as.numeric(as.factor(tec$index.i5))
```

### Load Biological table:

```{r}
bio <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/20221115_Datos Pacientes.xlsx")
bio <- bio[2:nrow(bio),]
bio <- bio[bio$Muestra %in% colnames(counts.tpm),]
```


### Filter lowely expressed genes:

We apply this filter here, then we will apply LIMMA. Keep genes that have more than 50% of non-zero counts across all samples:
```{r}
keep <- apply(counts.tpm, 1, function(x) sum(x == 0) <= 70)
#sum(keep) # 17,970

counts.tpm <- counts.tpm[keep,]

# Check how this filter affects cases and controls:
cases.tpm <- counts.tpm[,1:97]
controls.tpm <- counts.tpm[,98:ncol(counts.tpm)]

#sum(apply(cases.tpm, 1, function(x) sum(x == 0) <= 48)) # 17,884
#sum(apply(controls.tpm, 1, function(x) sum(x == 0) <= 22)) # 17,314
```

### Transform to DGE format:

```{r}
dge <- DGEList(counts=counts.tpm)
```

### Apply TMM Normalization:

```{r}
dge <- calcNormFactors(dge, method = "TMM")
```

### Apply logCPM transformation to test the effects of covariates:

```{r}
dgeTpmAdjCpm <- cpm(dge)
```

### Perform PCA:

```{r}
my_pca <- prcomp(t(dgeTpmAdjCpm), center = T, scale. = T)
summaryPca <- summary(my_pca)
```

### PCA Plots:

```{r, include=TRUE}
barplot((my_pca$sdev^2 / sum(my_pca$sdev^2))[1:20], names.arg = 1:20,  
        xlab = "PC #", ylab = "Amount of explained variance per PC",
        ylim=c(0,0.25), las=2, cex.names=.8)
```

```{r, include=TRUE}
plot(cumsum(my_pca$sdev^2 / sum(my_pca$sdev^2))[1:10], type="b", ylab = "Cummulative proportion of Variance Explained")
```

### PCA Plot:

```{r}
#png(file = "C://Users/Gerard/Desktop/AAA/RNAseq/Plots/PCA_Plot.png")
ggbiplot(my_pca, labels = colnames(dgeTpmAdjCpm), groups = as.factor(tec$Case), var.axes = F, obs.scale = 1, var.scale = 1) +
  scale_color_discrete(name = "") +
  theme(legend.direction = "horizontal", legend.position = "top")
#dev.off()
```



### Covariates against PCA:

#### Case:

```{r}
covariate <- "Case"
boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) cor.test(boxplotMatrix[,x],as.numeric(boxplotMatrix[,covariate]), method="spearman", exact = FALSE)[c(3,4)])), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "correlation")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$correlation <- as.numeric(format(dfPvalue$correlation, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Case)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- c()
Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### FC:

```{r}
covariate <- "FC"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=FC)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Lane:

```{r}
covariate <- "Lane"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Lane)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Library Date:

```{r}
covariate <- "Library"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Library)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### GC_Mean:

```{r}
covariate <- "GC_Mean"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=GC_Mean)) +
  geom_point() +
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### RIN:

```{r}
covariate <- "RIN"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=RIN)) +
  geom_point() +
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```


#### DV200:

```{r}
covariate <- "DV200"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=DV200)) +
  geom_point() +
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```


#### Qubit:

```{r}
covariate <- "Qubit"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Qubit)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### index.i5:

```{r}
covariate <- "index.i5"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=index.i5)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### index.i7:

```{r}
covariate <- "index.i7"

boxplotMatrix <- cbind(tec, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=index.i7)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```


#### Prepare and save Correlations table:

```{r}
rownames(Correlations) <- paste0("PC",1:10)
colnames(Correlations) <- c("CASE","FC","LANE","LIBRARY","GC_MEAN","RIN","DV200","QUBIT","INDEX.I5","INDEX.I7")

#write.xlsx(as.data.frame(Correlations), "C://Users/Gerard/Desktop/AAA/RNAseq/Corr_Tec_All.xlsx", sep = "\t", rowNames = T)
```


## Biological covariates against PCA:

### Sex:

```{r}
covariate <- "SEXO"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=SEXO)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- c()
Correlations <- cbind(Correlations, dfPvalue[,-1])
```

### Age:

```{r}
covariate <- "age"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=age)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```


### Type of aneurysm:

```{r}
covariate <- "TIPOANEURISMA"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=TIPOANEURISMA)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Height:

```{r}
covariate <- "Height_cm"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Height_cm)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

### Weight:

```{r}
covariate <- "Weight_Kg"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Weight_Kg)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

### Smoking:

```{r}
covariate <- "Smoking"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=Smoking)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

### Aortic Diameter:

```{r}
covariate <- "aortc_diameter_mm"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

mid <- mean(na.omit(boxplotMatrix[,covariate]))
ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=aortc_diameter_mm)) +
  geom_point()+
  scale_color_gradient2(midpoint=mid, low="darkblue", mid="lightgrey",
                        high="darkred", space ="Lab" ) +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```


### Other Aneurysms:

```{r}
covariate <- "other_aneurysm"

boxplotMatrix <- cbind(bio, my_pca$x[,1:10])
boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))
colnames(boxplotMatrixMelt)[1] <- "covariate"

dfPvalue <- as.data.frame(matrix(unlist(lapply(paste0("PC", 1:10), function(x) c(anova(lm(as.formula(paste0(x, "~1+", covariate)), data=boxplotMatrix))[5][[1]][1], summary(lm(as.formula(paste0(x, "~1+", covariate)), data = boxplotMatrix))[9]))), ncol=2, byrow = T))

colnames(dfPvalue) <- c("pvalue", "adjR2")
rownames(dfPvalue) <- paste0("PC", 1:10)

dfPvalue$pvalue <- as.numeric(format(dfPvalue$pvalue, digits = 3))
dfPvalue$adjR2 <- as.numeric(format(dfPvalue$adjR2, digits = 3))

ggplot(boxplotMatrix, aes(x=PC1, y=PC2, color=other_aneurysm)) +
  geom_point() +
  scale_color_discrete() +
  theme_classic()
```

```{r}
colnames(dfPvalue) <- c("pvalue", "covariate")
dfPvalue$covariate <- ifelse(dfPvalue$pvalue < 0.05, paste0(dfPvalue$covariate, " ***"), dfPvalue$covariate)

Correlations <- cbind(Correlations, dfPvalue[,-1])
```

#### Prepare and save Correlations table:

```{r}
rownames(Correlations) <- paste0("PC",1:10)
colnames(Correlations) <- c("SEX","AGE","TYPE","HEIGHT","WEIGHT","SMOKING","DIAMETER","O.ANEURYSM")

write.xlsx(as.data.frame(Correlations), "C://Users/Gerard/Desktop/AAA/RNAseq/Corr_Bio_All.xlsx", sep = "\t", rowNames = T)
```


1. Recommended Technical Variables: FC, LANE, GC_MEAN, RIN, DV200, QUBIT. (LIBRARY not included because is extremely correlated with Case)

2. Recommended Biological Variables: SEX, AGE. With and Without SMOKING.

