---
title: "Limma RNAseq AAA"
author: "Gerard Temprano Sagrera"
date: "23/11/2022"
output: html_document
---

```{r, warning=FALSE, message=FALSE, echo = FALSE}
library(limma);library(data.table);library(edgeR);library(tidyverse);library(openxlsx)
```

```{r}
#limmaUsersGuide()
```

### Make a Count Matrix from the RSEM results:

```{r}
ff <- list.files("/home/gerard/RNAseq/GTEx",pattern = "*.rsem.genes.results", recursive = T, full.names = T)

count.matrix <- c()

# Loop through all *.rsem.genes.results files obtained from RSEM.
for (i in 1:length(ff)) {
  if (i == 1) {
    res <- fread(ff[i], select = c("gene_id", "TPM"))
    sample <- strsplit(ff[i], "/")[[1]][6]
    names(res)[names(res) == "TPM"] <- sample
    count.matrix <- cbind(count.matrix, res)
  } else {
    res <- fread(ff[i], select = c("TPM"))
    sample <- strsplit(ff[i], "/")[[1]][6]
    names(res)[names(res) == "TPM"] <- sample
    count.matrix <- cbind(count.matrix, res)
  }
}

if (ncol(count.matrix) != 142) {warning("Not all samples have been analyzed")}

count.matrix <- count.matrix %>% remove_rownames %>% column_to_rownames(var="gene_id")
```


```{r}
# Use the annotation file to select among the quantified genes:
ann <- fread("C://Users/Gerard/Desktop/AAA/RNAseq/Annotation/gencode.v26.annotation.fixed.gtf.gz")
count.matrix2 <- count.matrix[rownames(count.matrix) %in% ann$gene_id,]

paste("From the", nrow(ann), "filtered genes present in the annotation file,", nrow(count.matrix2), "have been quantified")

paste(nrow(count.matrix) - nrow(count.matrix2), "genes were quantified but are removed due to filtering in the annotation file")

if (sum(duplicated(rownames(count.matrix2))) != 0) {warning("There are duplicates gene ids!!")}
```

### Convert to DGElist object and remove rows with consistent 0 or low counts:

```{r}
dge <- DGEList(counts=count.matrix2)

# Set "Control" to be the first level
group_list <- factor(x = c(rep("Case",5), rep("Control",4)),
              levels=c("Control","Case"))

design <- model.matrix(~group_list)

keep <- filterByExpr(dge, design)
dge <- dge[keep,,keep.lib.sizes=FALSE]

paste("We exclude", sum(!keep), "genes and we have", sum(keep), "genes that pass all our filters")
```

### Apply TMM Normalization:

```{r}
dge <- calcNormFactors(dge, method = "TMM")
max(dge$samples$lib.size)/min(dge$samples$lib.size)
```

```{r}
dgeTpmAdjCpm <- cpm(dge)
```

### Perform PCA:

```{r}
my_pca <- prcomp(t(dgeTpmAdjCpm), center = TRUE,scale. = TRUE)
summaryPca <- summary(my_pca)

saveRDS(my_pca,"/home/gerard/PCA.rds")
```

```{r}
# To run it locally:
my_pca <- readRDS("C://Users/Gerard/Downloads/PCA.rds")
```

```{r, include=TRUE}
# Cumulative variance plot:
plot(cumsum(my_pca$sdev^2 / sum(my_pca$sdev^2)), type="b", ylab = "Cummulative proportion of Variance Explained")
```

```{r, include=TRUE}
barplot((my_pca$sdev^2 / sum(my_pca$sdev^2))[1:20], names.arg = 1:20,  
        xlab = "PC #", ylab = "Amount of explained variance per PC",
        ylim=c(0,0.5), las=2, cex.names=.8)
```

Correlation of the most important PCs (PC1-4) with our technical:

-   FC

-   Date

-   Batch

-   GC_Mean

-   RIN

-   DV200

-   Qubit

and biological covariates:

-   Age

-   Sex


```{r}
#tec <- read.xlsx("/home/gerard/RNAseq/Technical_CompleteV2.xlsx")
tec <- read.xlsx("C://Users/Gerard/Desktop/AAA/RNAseq/Tables/Technical_CompleteV2.xlsx")
names(tec)
```


```{r}
ss <- c("22ID00809","22ID00810","22ID00811","22ID00812","22ID00813","22ID00906","22ID00907","22ID00908","22ID00909")
tec2 <- tec[tec$Seq_ID %in% ss,]
```


```{r}
boxplotMatrix <- cbind(tec2, my_pca$x[,1:4])

boxplotMatrixMelt <- data.frame(reshape2::melt(boxplotMatrix[,c(covariate, paste0("PC", 1:4))], id.vars=covariate))

colnames(boxplotMatrixMelt)[1] <- "covariate"
```






### Using normalized values:

```{r}
v.norm <- voom(dge, design, plot=TRUE)
fit <- lmFit(v.norm,design)
fit2 <- eBayes(fit)

tempOutput.norm <- topTable(
    fit2, n=Inf,
    adjust.method = 'BH',
    coef='group_listCase')
```


### Using matrix of counts directly:

```{r}
v.counts <- voom(count.matrix, design, plot=TRUE)
fit <- lmFit(v.counts,design)
fit2 <- eBayes(fit)

tempOutput.counts <- topTable(
    fit2, n=Inf,
    adjust.method = 'BH',
    coef='group_listCase')
```

### Using matrix of counts and normalize:

```{r}
v <- voom(count.matrix,design,normalize="quantile", plot = T)
fit <- lmFit(v,design)
fit2 <- eBayes(fit)

tempOutput <- topTable(
    fit2, n=Inf,
    adjust.method = 'BH',
    coef='group_listCase')
```

```{r}
tempOutput.norm
```

```{r}
tempOutput.counts
```

```{r}
tempOutput
```

```{r}
summary(decideTests(fit[,-1]))
```
